Efficient Data Provenance Analysis for Concurrent Programs

### 2013-11-13 paper.tex
committing before lunch.  need to do a read through.  Then, need to add some more debugging data points.  Is there time?

\section{Evaluation}
\label{sec:eval}
We evaluated last writer slices and \ctraps along several axes.  We show that
last writer slices provide information that is useful during debugging in
Section~\ref{sec:eval:debugging} looking at a set of buggy programs and
comparing to prior, related work.  We evaluate our performance overheads and
show that last writer slices and \ctraps are feasible for use in deployed
systems for a majority of workloads we studied.  We then characterize our
instrumentation and discuss the sources of our system's run time overhead.   We
show that \ctraps is useful by implementing two existing applications and
analyzing their performance.

@@

%For the remaining case, {\tt apache2}, last writer slices would
%probably help with debugging, but it is less obvious because this bug is a
%complex, multi-variable bug.  We provide more detail on {\tt apache2} in
%Section~\ref{sec:eval:debugging:cases}.

Our evaluation shows that bad value origin tracking~\cite{badapples} is not
helpful in three out of five cases where last writer slices is
useful.  In one case, a use-after-delete bug in {\tt pbzip2}, bad value origin
tracking helps.  For this bug, when one thread deletes a variable, it makes
that variable unusable and bad value origin tracking would track that value.
When the failure occurs, the programmer would then know why the value was
deleted.

In the remaining case, {\tt apache}, bad value origin tracking may help.  The
bug is an atomicity violation that leads to a double free.  Two threads
manipulate an object's reference count and deallocate the object when the count
hits zero. Zero-checks and deallocation should be atomic, but the code does not
ensure they are.  If threads interleave their code, multiple threads might see
zero reference counts, which is incorrect.  When that happens, both threads try
to free the object and the second free causes a crash.  Bad value origin
tracking may help in this case.  At the first free, the freed pointer becomes
unusable and its origin is tracked.  At the crash, the bad value's reported
origin is the code point that frees the object.  That information potentially
helps debug the program by suggesting a double free.  However, in {\tt apache},
there are many threads, some perform different tasks, and all may free that
object at that code point.  Bad value origin tracking does not record the
thread of the origin.  The programmer is left unsure of whether the problem is
with single-threaded logic or with synchronization.  In contrast, last writer
slices reports the freed value's origin's thread and code point, making it
clear that the programmer is dealing with a concurrency bug, and which threads
are involved.

\subsubsection{Case Studies}
\label{sec:eval:debugging:cases}

@@

\paragraph{JDK-StringBuffer}
The running example in Figures~\ref{fig:coreDumpFail} and ~\ref{fig:jdklws},
and the text in Section~\ref{sec:debugging} describes the benefit provided by
last writer slices in debugging an ordering violation in the JDK-1.4.

@@

\paragraph{HTTrack}
Figure~\ref{fig:httlws} shows how last writer slices help understand the root
cause of a concurrency bug from HTTrack-3.43.9.  The program crashes at point
{\bf B} when it calls {\tt hts\_mutex\_lock()} with an invalid mutex.    The
root cause of this bug is that Thread 1 fails to initialize the mutex before
Thread 2 uses it.   The core dump is of little debugging use, only showing
that, indeed, the lock is uninitialized.  Even if the programmer compared the
core dump from the failing run to the memory image from a correct execution --
a helpful debugging strategy -- they would find only that in a correct
execution, the lock contains valid mutex state.

@@

\paragraph{MySQL}

>>>>

\section{Evaluation}
\label{sec:eval}
We evaluated last writer slices and \ctraps along several axes.  We show that
last writer slices provide information that is useful during debugging in
Section~\ref{sec:eval:debugging} looking at a set of buggy programs and
comparing to prior work.  We evaluate our performance overheads and
show that last writer slices and \ctraps are feasible for use in deployed
systems for a majority of workloads we studied.  We then characterize our
instrumentation and discuss sources run time overhead.   We
show that \ctraps is useful by implementing two existing applications and
analyzing their performance.

@@

%For the remaining case, {\tt apache2}, last writer slices would
%probably help with debugging, but it is less obvious because this bug is a
%complex, multi-variable bug.  We provide more detail on {\tt apache2} in
%Section~\ref{sec:eval:debugging:cases}.
Our evaluation shows that bad value origin tracking~\cite{badapples} is not
helpful in three out of five cases where last writer slices is useful.  
There are two cases where bad value origin tracking may help.  For a
use-after-delete bug in {\tt pbzip2}, bad value origin tracking helps.  For
this bug, when one thread deletes a variable, it makes that variable unusable
and bad value origin tracking would track that value.  When the failure occurs,
the programmer would then know why the value was deleted.

For {\tt apache}, bad value origin tracking may help.  The bug is an atomicity
violation that leads to a double free.  Two threads manipulate an object's
reference count and deallocate the object when the count hits zero. Zero-checks
and deallocation should be atomic, but the code does not ensure they are.  If
threads interleave their code, multiple threads might see zero reference
counts, which is incorrect.  When that happens, both threads try to free the
object and the second free causes a crash.

Bad value origin tracking may help in this case.  At the first free, the freed
pointer becomes unusable and its origin is tracked.  At the crash, the bad
value's reported origin is the code point of the free. That information may
help debug the program by suggesting a double free.  However, in {\tt apache},
many threads perform different tasks and all call the same free code.  Bad
value origin tracking does not record the thread information.  The programmer
is left unsure whether the problem is with single-threaded logic or
synchronization.  In contrast, last writer slices reports the code point and
thread, making it clear the bug is a concurrency bug and identifying the
involved threads.

\subsubsection{Debugging Case Studies}
\label{sec:eval:debugging:cases}

@@

\paragraph{jdk-1.4}
The running example in Figures~\ref{fig:coreDumpFail} and ~\ref{fig:jdklws},
and the text in Section~\ref{sec:debugging} describes the benefit provided by
last writer slices in debugging an ordering violation in the JDK-1.4.

@@

\paragraph{httrack}
Figure~\ref{fig:httlws} shows how last writer slices help understand the root
cause of a concurrency bug from HTTrack-3.43.9.  The program crashes at point
{\bf B} when it calls {\tt hts\_mutex\_lock()} with an invalid mutex.    The
root cause of this bug is that Thread 1 fails to initialize the mutex before
Thread 2 uses it.   The core dump is of little debugging use, only showing
that, indeed, the lock is uninitialized.  Even if the programmer compared the
core dump from the failing run to the memory image from a correct execution --
a helpful debugging strategy -- they would find only that in a correct
execution, the lock contains valid mutex state.

@@

\paragraph{mysql}
